### 2018-05-17 Report/notes.tex
last modif

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% LaTeX Example: Project Report
%
% Source: http://www.howtotex.com
%
% Feel free to distribute this example, but please keep the referral
% to howtotex.com
% Date: March 2011 
% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% How to use writeLaTeX: 
%
% You edit the source code here on the left, and the preview on the
% right shows you the result within a few seconds.
%
% Bookmark this page and share the URL with your co-authors. They can
% edit at the same time!
%
% You can upload figures, bibliographies, custom classes and
% styles using the files menu.
%
% If you're new to LaTeX, the wikibook is a great place to start:
% http://en.wikibooks.org/wiki/LaTeX
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Edit the title below to update the display in My Documents
%\title{Project Report}
%
%%% Preamble
\documentclass[paper=a4, fontsize=11pt]{scrartcl}
\usepackage[T1]{fontenc}
\usepackage{fourier}

\usepackage[english]{babel}															% English language/hyphenation
\usepackage[protrusion=true,expansion=true]{microtype}	
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage[pdftex]{graphicx}	
\usepackage{url}

%%% Custom sectioning
\usepackage{sectsty}
\allsectionsfont{\centering \normalfont\scshape}
\usepackage{tikz}
\usetikzlibrary{chains,shapes.multipart}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}

%%% Custom headers/footers (fancyhdr package)
\usepackage{fancyhdr}
\pagestyle{fancyplain}
\fancyhead{}											% No page header
\fancyfoot[L]{}											% Empty 
\fancyfoot[C]{}											% Empty
\fancyfoot[R]{\thepage}									% Pagenumbering
\renewcommand{\headrulewidth}{0pt}			% Remove header underlines
\renewcommand{\footrulewidth}{0pt}				% Remove footer underlines
\setlength{\headheight}{13.6pt}

%%% Equation and float numbering
\numberwithin{equation}{section}		% Equationnumbering: section.eq#
\numberwithin{figure}{section}			% Figurenumbering: section.fig#
\numberwithin{table}{section}				% Tablenumbering: section.tab#

%%% Maketitle metadata
\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} 	% Horizontal rule

\title{
	%\vspace{-1in} 	
	\usefont{OT1}{bch}{b}{n}
	\normalfont \normalsize \textsc{School of random department names} \\ [25pt]
	\horrule{0.5pt} \\[0.4cm]
	\huge This is the title of the template report \\
	\horrule{2pt} \\[0.5cm]
}
\author{
	\normalfont 								\normalsize
	Firstname Lastname\\[-3pt]		\normalsize
	\today
}
\date{}

\usepackage{listings}
\usepackage{color}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\definecolor{prismgreen}{rgb}{0, 0.6, 0}

%% BEGIN PRISM CODE LISTING 
\lstdefinelanguage{Prism}{ % syntax highlight via font
	basicstyle=\color{red}\tiny\ttfamily, % small true type font (like
	courier)
	morekeywords={bool,C,ceil,const,ctmc,double,dtmc,endinit,endmodule,endrewards,endsystem,F,false,floor,formula,G,global,I,init,int,label,max,mdp,min,module,nondeterministic,P,Pmin,Pmax,prob,probabilistic,R,rate,rewards,Rmin,Rmax,S,stochastic,system,true,U,X},
	keywordstyle={\bfseries\color{black}},
	numberstyle=\tiny\color{black},
	comment=[l] {//}, morecomment=[s]{/*}{*/}, % single and multi-line
	commentstyle= \color{prismgreen}, % dark green
	tabsize=4, % tab treatment (going to be fixed in Prism)
	captionpos=b, % put captions at the bottom
	escapechar=@ % write LaTeX comments escaped by @ symbol
}

%define command \prism with one argument for inline printing of \prismcode
\newcommand{\prism}[1]{\lstinline[language=Prism,basicstyle=\small
	\ttfamily]|#1|}

%% END PRISM CODE LISTING

     % Set your language (you can change the language for each code-block optionally)

%%% Begin document
\begin{document}
	\maketitle
	\begin{abstract}
		The subject of this internship is first modeling some system  in the PrismGSMP. That is a continuous system with rates instead of probabilistic transition. Transitiom are syncronized with event which follows some probabilistic laws like exponential law or dirac. One good property about the exponenial law is the events are markovian. That is the probability of the event of happening doenst depends on the time already spend.

	\end{abstract}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%SECTION1%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	\section{Queue : A toy example}
	We considere a M/M/1/n queue (see Kendall notation).

	\subsection{first model}

	\begin{figure}
		\centering
		\begin{tikzpicture}[start chain=going right,>=latex,node distance=0pt]
		% the rectangular shape with vertical lines
		\node[rectangle split, rectangle split parts=6,
		draw, rectangle split horizontal,text height=1cm,text depth=0.5cm,on chain,inner ysep=0pt] (wa) {};
		\fill[white] ([xshift=-\pgflinewidth,yshift=-\pgflinewidth]wa.north west) rectangle ([xshift=-15pt,yshift=\pgflinewidth]wa.south);

		% the circle
		\node[draw,circle,on chain,minimum size=1.5cm] (se) {$\mu$};

		% the arrows and labels
		\draw[->] (se.east) -- +(20pt,0);
		\draw[<-] (wa.west) -- +(-20pt,0) node[left] {$\lambda$};
		\node[align=center,below] at (wa.south) {Producing \\ Area};
		\node[align=center,below] at (se.south) {Consume};
		\end{tikzpicture}
		\caption{a queue}
		\label{fig:queue0}
	\end{figure}

	\begin{figure}
		\centering
		\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.8cm,
		semithick]
		\tikzstyle{every state}=[fill=red,draw=none,text=white]

		\node[initial,state] (s0)                    {$0$};
		\node[state]         (s1) [right of=s0] {$1$};
		\node[state]         (s2) [right of=s1] {$2$};
		\node[state]         (s3) [right of=s2] {$3$};
		\node[state]         (s4) [right of=s3]       {$4$};
		\node[state]         (sw) [right of=s4,fill=white,text=black]       {...};
		\node[state]         (sn) [right of=sw]       {$n$};

		\path (s0) edge [bend left]             node {Dirac(t)} (s1)
		(s1) edge [bend left] node {$\lambda$} (s0)
		edge  [bend left]            node {Dirac(t)} (s2)
		(s2) edge [bend left]             node {$\lambda$} (s1)
		edge [bend left]  node {Dirac(t)} (s3)
		(s3) edge [bend left]              node {$\lambda$} (s2)
		edge [bend left]  node {Dirac(t)} (s4)
		(s4) edge [bend left]             node {$\lambda$} (s3)
		edge [bend left]  node {Dirac(t)} (sw)
		(sw) edge [bend left]             node {$\lambda$} (s4)
		edge [bend left]  node {Dirac(t)} (sn)
		(sn) edge [bend left]             node {$\lambda$} (sw);
		\end{tikzpicture}
		\caption{Queue with n capacity, the current size is written on the states}
		\label{fig:Queue1}
	\end{figure}

	\begin{figure}
		\centering
		\lstinputlisting[language=Prism]{../Model/timeoutqueue.sm}

		\caption{Prism code for figure \ref{fig:Queue1}}
		\label{fig:queuecode1}
	\end{figure}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%SECTION2%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	\subsection{second model(using phase type fitting)}
	The timeout part is replaced by $k$ $\mu$ transition following the exponential law of parameter $\mu$. Here on the figure, k=5, which mean that you need 5 consecutive $\mu$ transition to $produce$.

	\begin{figure}
		\centering
		\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.0cm,
		semithick]
		\tikzstyle{every state}=[fill=red,draw=none,text=white]

		\node[state] (s0)                    {$i$};
		\node[state] (s02) [right of=s0,fill=blue,scale=0.5]                   {};
		\node[state] (s03) [right of=s02,fill=blue,scale=0.5]                    {};
		\node[state] (s04) [right of=s03,fill=blue,scale=0.5]                   {};
		\node[state] (s05) [right of=s04,fill=blue,scale=0.5]                    {};
		\node[state] (s1) [below of=s0]                   {$i+1$};

		\node[state] (s12) [right of=s1,fill=blue,scale=0.5]                   {};
		\node[state] (s13) [right of=s12,fill=blue,scale=0.5]                    {};
		\node[state] (s14) [right of=s13,fill=blue,scale=0.5]                   {};
		\node[state] (s15) [right of=s14,fill=blue,scale=0.5]                    {};
		\node[state] (s2) [below of=s1]     {$i+2$};

		\node[state] (sw0) [above of=s0,fill=white,text=black]        {...}            {};
		\node[state] (sw1) [right of=s2,fill=white,text=black]       {...}             {};

		\path (s0) edge   node {$\lambda$} (sw0)
		edge   node {$\mu$} (s02)

		(s02)   edge   node {$\lambda$} (sw0)
		edge   node {$\mu$} (s03)
		(s03)   edge  node {$\lambda$} (sw0)
		edge   node {$\mu$} (s04)
		(s04) edge  node {$\lambda$} (sw0)
		edge   node {$\mu$} (s05)
		(s05) edge  node {$\lambda$} (sw0)
		edge   node {$\mu$} (s1)
		(s1)    edge  node {$\lambda$} (s0)
		edge   node {$\mu$} (s12)
		(s12)   edge  node {$\lambda$} (s0)
		edge   node {$\mu$} (s13)
		(s13) edge  node {$\lambda$} (s0)
		edge   node {$\mu$} (s14)
		(s14) edge  node {$\lambda$} (s0)
		edge   node {$\mu$} (s15)
		(s15) edge  node {$\lambda$} (s0)
		edge   node {$\mu$} (s2)
		(s2) edge  node {$\lambda$} (s1)
		edge   node {$\mu$} (sw1)    
		(sw1) edge  node {$\lambda$} (s1);
		\end{tikzpicture}
		\caption{Queue with phase type fitting of parameter k}
		\label{fig:queueptf1}
	\end{figure}

	\begin{figure}
		\centering
		\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.0cm,
		semithick]
		\tikzstyle{every state}=[fill=red,draw=none,text=white]

		\node[state] (s0)                    {$i$};
		\node[state] (s02) [right of=s0,fill=blue,scale=0.5]                   {};
		\node[state] (s03) [right of=s02,fill=blue,scale=0.5]                    {};
		\node[state] (s04) [right of=s03,fill=blue,scale=0.5]                   {};
		\node[state] (s05) [right of=s04,fill=blue,scale=0.5]                    {};
		\node[state] (s1) [below of=s0]                   {$i+1$};

		\node[state] (s12) [right of=s1,fill=blue,scale=0.5]                   {};
		\node[state] (s13) [right of=s12,fill=blue,scale=0.5]                    {};
		\node[state] (s14) [right of=s13,fill=blue,scale=0.5]                   {};
		\node[state] (s15) [right of=s14,fill=blue,scale=0.5]                    {};
		\node[state] (s2) [below of=s1]     {$i+2$};

		\node[state] (sw0) [above of=s0,fill=white,text=black]        {...}            {};
		\node[state] (sw02) [right of=sw0,fill=white,text=black]        {...}            {};
		\node[state] (sw03) [right of=sw02,fill=white,text=black]        {...}            {};
		\node[state] (sw04) [right of=sw03,fill=white,text=black]        {...}            {};
		\node[state] (sw05) [right of=sw04,fill=white,text=black]        {...}            {};
		\node[state] (sw1) [right of=s2,fill=white,text=black]       {...}             {};

		\path (s0) edge   node {$\lambda$} (sw0)
		edge   node {$\mu$} (s02)

		(s02)   edge   node {$\lambda$} (sw02)
		edge   node {$\mu$} (s03)
		(s03)   edge  node {$\lambda$} (sw03)
		edge   node {$\mu$} (s04)
		(s04) edge  node {$\lambda$} (sw04)
		edge   node {$\mu$} (s05)
		(s05) edge  node {$\lambda$} (sw05)
		edge   node {$\mu$} (s1)
		(s1)    edge  node {$\lambda$} (s0)
		edge   node {$\mu$} (s12)
		(s12)   edge  node {$\lambda$} (s02)
		edge   node {$\mu$} (s13)
		(s13) edge  node {$\lambda$} (s03)
		edge   node {$\mu$} (s14)
		(s14) edge  node {$\lambda$} (s04)
		edge   node {$\mu$} (s15)
		(s15) edge  node {$\lambda$} (s05)
		edge   node {$\mu$} (s2)
		(s2) edge  node {$\lambda$} (s1)
		edge   node {$\mu$} (sw1)    
		(sw1) edge  node {$\lambda$} (s12);
		\end{tikzpicture}
		\caption{Queue with phase type fitting of parameter k}
		\label{fig:queueptf2}
	\end{figure}

	\begin{figure}
		\centering
		\begin{lstlisting}
		gsmp

		const int k = 10;
		const int qCapacity = 10;

		const double timeout = 10;
		const double lambda = 0.2;

		module main

		qSize : [0..qCapacity] init 0;

		[produce] (qSize < qCapacity) -> (qSize' = qSize+1);
		[consume] (qSize > 0) -> lambda: (qSize' = qSize - 1);

		endmodule

		module trigger

		i : [1..k+1];

		[] i < k -> k/timeout : (i'=i+1);
		[produce] i = k -> k/timeout : (i'=1);
		//[consume] true -> (i'=1);

		endmodule

		\end{lstlisting}
		\caption{Caption}
		\label{fig:queuecode2}
	\end{figure}

	\subsection{finding $\mu$}

	According to Prism documentation on phase type fittimg, we should take $\mu = t/k$. The mean of diract(t) is t, and the mean of exponential($\mu$) is $1/\mu$. Hence $\mu = t/k$.

	\subsection{Results}

	For a queue capacity of 10, we compare the steady state probabilities according to k the parameter of phase type fitting.

	\begin{figure}
		\centering
		\includegraphics[width=18cm]{sspqueue.png}
		\caption{The blue line is the result without phase type fitting}
		\label{fig:res1}
	\end{figure}

	\begin{figure}
		\centering
		\includegraphics[width=16cm]{time.png}
		\caption{time of computation versus k}
		\label{fig:time1}
	\end{figure}

	\begin{figure}
		\centering
		\includegraphics[width=16cm]{size.png}
		\caption{size of model versus k}
		\label{fig:time1}
	\end{figure}

	%%% End document
\end{document}\\