### 2018-07-23 Report/report/paper.tex
modification of my institution paper

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads,custombib]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

%my package
\usepackage[backend=biber,
%this is the label style in the bibliography at the end of document
style=numeric,
%this is the label style for each citation in the document
citestyle=numeric,
sorting=none,
%sortcae=false                  
%citestyle=authoryear
]{biblatex}
%\usepackage[backend=biber]{biblatex}
\addbibresource{paper.bib} %Imports bibliography file

\usepackage{amsmath}
\usepackage{amsfonts}

\usepackage{tikz}
\usetikzlibrary{chains,shapes.multipart}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}

\usepackage{listings}
\usepackage{color}
\usepackage{float}

%% BEGIN PRISM CODE LISTING 
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}

\lstset{frame=tb,
	language=Java,
	aboveskip=1mm,
	belowskip=1mm,
	showstringspaces=false,
	columns=flexible,
	basicstyle={\fontsize{8.8}{8.5}\ttfamily},
	numbers=none,
	numberstyle=\tiny\color{gray},
	keywordstyle=\color{blue},
	commentstyle=\color{dkgreen},
	breaklines=true,
	breakatwhitespace=true,
	tabsize=3,
	morekeywords={ctmc,event, gsmp, dirac, exponential, uniform, weibull, erlang,
		module, init, rewards, endrewards, label, true, endmodule}
}
\usepackage{csquotes}

\usepackage{hyperref}

\begin{document}
%
\title{Continuous-Time Markov Chain with Alarms analysis and comparaison of tools \thanks{Supported by organization x.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Thomas Mari\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{ENS-Rennes, Campus de Ker lann, Avenue Robert Schuman, 35170 Bruz, France \and
Masaryk University, Botanická 68A, 602 00 Brno-Královo Pole, Czech republic
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
	Phase-type fitting remains the only way of modeling non-Markovian distributions within PRISM model checker.
	When analyzing models with phase-type distributions, it is problematic to obtain results of sufficient precision within reasonable time.
	We ran experimental computations and deduced a reliable way of obtaining precise analysis results for phase-type fitted PRISM CTMCs.
	Lastly, we hint at an entirely different (and arguably better) approach to handling non-Markovian distributions.

%The abstract should briefly summarize the contents of the paper in
%150--250 words.

\keywords{PRISM model checker \and CTMC \and deterministic timeout \and phase-type distribution \and modeling \and analysis.}
\end{abstract}
%
%
%
\section{Introduction}

	PRISM \cite{KNP11} and Storm \cite{DBLP:journals/corr/DehnertJK017} are popular tools for modeling and analysis of stochastic systems in continuous time. They use efficient algorithms for analysis of continuous time Markov chains (CTMC). This approach suffers from a severe restriction of expressiveness, because the time between transitions must be exponentially distributed. This restriction can be remedied by the use of phase-type distributions, which can approximate any general distribution with arbitrary accuracy by only using exponential distributions \cite{Buchholz:2014:IMP:2683922}. However, the use of phase-type distributions drastically increases the number of states within the CTMC.

	In this paper, we experimentally evaluate the precision of the result and required computation time of various approaches to analysis of continuous time stochastic models with deterministic transitions (timeouts). The obtained results are then compared against results of our extension of PRISM capable of efficiently analyzing CTMC with non-Markovian alarms (ACTMC).

%	BIGGER INTRO

	We will go through the theoretical bases needed to understand the contribution. CITE LUBOSH'THESIS

\section{Related Work}
\input{related_work.tex}

\section{Preliminaries}
\input{preliminaries.tex}

	\section{Experimental evaluation}
	\label{experimetal_evaluation}

	\begin{figure}[H]
		\centering
		\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.67cm,
		semithick]
		\tikzstyle{every state}=[fill=white,draw=black,text=black]

		\node[initial, initial text ={},state] (s0)                    {$0$};
		\node[state]         (s1) [right of=s0] {$1$};
		\node[state]         (s2) [right of=s1] {$2$};
		\node[state]         (s3) [right of=s2] {$3$};
		\node[state]         (s4) [right of=s3]       {$4$};
		\node[state]         (sw) [right of=s4,fill=white,text=black]       {...};
		\node[state]         (sn) [right of=sw]       {$10$};

		\path[->, dashed]  
		(s0) edge [bend left]             node {$t$} (s1);
		\path[->]
		(s1) edge [bend left] node {$\lambda$} (s0);
		\path[->, dashed]
		(s1) edge  [bend left]            node {$t$} (s2);
		\path[->]
		(s2) edge [bend left]             node {$\lambda$} (s1);
		\path[->, dashed]
		(s2) edge [bend left]  node {$t$} (s3);
		\path[->]
		(s3) edge [bend left]              node {$\lambda$} (s2);
		\path[->, dashed]
		(s3) edge [bend left]  node {$t$} (s4);
		\path[->]
		(s4) edge [bend left]             node {$\lambda$} (s3);
		\path[->, dashed]
		(s4) edge [bend left]  node {$t$} (sw);
		\path[->]
		(sw) edge [bend left]             node {$\lambda$} (s4);
		\path[->, dashed]
		(sw) edge [bend left]  node {$t$} (sn);
		\path[->]
		(sn) edge [bend left]             node {$\lambda$} (sw);
		\path[->, dashed]
		(sn) edge [loop right]  node {$t$} (sn);
		\end{tikzpicture}
		\caption{Model of a D/M/1/10 queue with production timeout $t = 0.1$ and service rate $\lambda = \frac{ln(2)}{t} \approx 6.931471805599453094$. The deterministic arrival event does not reset when the exponential service event occurs.}
		\label{fig:D/M/1/10}
	\end{figure}

	We have chosen to model a D/M/1/10 queue (deterministic arrivals, exponentially distributed service, single server, capacity of 10) as shown in Figure \ref{fig:D/M/1/10}. The deterministic arrivals are approximated by a phase-type distribution, so the model remains a CTMC and can be analyzed by PRISM. The PRISM model we used is shown in Figure \ref{fig:D/M/1/10_PRISM}.
	We computed the steady-state distribution with various values of PRISM termination epsilon $\epsilon$ and the number of phases $k$.

	\begin{figure}[H]
		\begin{lstlisting}
		ctmc

		const int k;
		const int qCapacity = 10;
		const double timeout = 0.1;
		const double lambda = 6.931471805599453094;

		module main

		qSize : [0..qCapacity] init 0;

		[produce] (qSize <= qCapacity) -> (qSize' = min(qSize+1,qCapacity));
		[consume] (qSize > 0) -> lambda: (qSize' = qSize - 1);

		endmodule

		module trigger

		i : [1..k+1];

		[] i < k -> k/timeout : (i'=i+1);
		[produce] i = k -> k/timeout : (i'=1);

		endmodule
		\end{lstlisting}
		\caption{PRISM CTMC model of the D/M/1/10 queue as shown in Figure \ref{fig:D/M/1/10} with a phase-type distribution approximating the arrival timeout. The phase-type module \emph{trigger} is used as suggested on the \href{http://www.prismmodelchecker.org/manual/FrequentlyAskedQuestions/PRISMModelling}{PRISM website}. Phase-type parameter $k$ represents the number of phases. Increasing $k$ should improve the approximation of the timeout. }
		\label{fig:D/M/1/10_PRISM}
	\end{figure}

	%\pagebreak
	\subsection{Obtaining precise results for reference}
	\label{reference}

	%%%	In order to obtain some steady-state probabilities of reference to compare precision of different tools and engines, we computed the same steady-state probabilities two different ways. The first one is the phase type fitting method described as above (ref to the first page). The second one is by using our PRISM ACTMC implementation with precise termination epsilon.
	%%%	 the computation  results are quite similar as shown in picture citation

	%%%	 put picture diff_per_stqte 
	%%%	 explain picture

	%%% little conclusion: we can from now trust the event model as a reference

	%%	 In order to compare the precision of the results obtained by the experiments, a trustworthy high-precision reference is needed. To achieve this, we used our own PRISM ACTMC implementation, which is capable of analyzing the given model directly without phase-type fitting. Using our implementation, we have computed the steady-state probabilities for very precise termination epsilon $\epsilon = 10^{-20}$ and compared it to the steady-state probabilities obtained by PRISM CTMC with phase-type fitting for $\epsilon = 10^{-16}$ and $k = 1000000$. As shown in Figure \ref{fig:diff_storm_dynamic}, the differences between the computations are smaller than $10^6$. Under the hypothesis that the true result is closer to the result obtained from PRISM ACTMC than to the result obtained from Storm, we can conclude that the PRISM ACTMC can serve as reference with precision at least $10^{-6}$.%%

	In order to compare the precision of the results obtained by the experiments, a trustworthy high-precision reference is needed. To achieve this, we have modeled the given D/M/1/10 queue in Oris \cite{Bucci2010}. The steady-state distribution obtained by Oris equals the steady-state distribution obtained by our PRISM ACTMC implementation up to at least 14 digits. Hence, our tool can be considered a trustworthy reference with precision of $10^{-14}$.

	I THINK AT THIS POINT OF THE ARTICLE, IT IS NOT OBVIOUS THAT OUR TOOL REFER TO PRISM GSMP 
	%\pagebreak
	\subsection{Default absolute termination epsilon $\epsilon = 10^{-6}$}
	\label{default_absolute}

	First, we used the PRISM default absolute termination epsilon $\epsilon = 10^{-6}$. The results are shown in Figure \ref{fig:distance_constant_standard_constant_kappa}. For this $\epsilon$, the obtained results are arguably insufficient. Increasing $k$ does little to improve the precision, and only up to about $k \leq 50$, at which point increasing $k$ further starts making the precision worse.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[trim=0.3cm 0cm 0cm 1.35cm,width=15cm, clip]{../picture/New_model/1E-6/distance_constant_1E-6.png}
		\end{center}
		\caption{Maximum distance (infinity norm) of the steady-state probabilities from the reference high-precision ACTMC ($10^{-20}$) against various values of $k$ for default absolute termination epsilon $\epsilon = 10^{-6}$. ACTMC results for $\epsilon = 10^{-6}$ are also compared to the reference.}
		\label{fig:distance_constant_standard_constant_kappa}
	\end{figure}

	%\pagebreak
	\subsection{Lowering absolute termination epsilon to $\epsilon = 10^{-10}$}
	\label{lowering}

	For much lower $\epsilon = 10^{-10}$, the precision of the result improves. However, for higher $k$ the precision continues to deteriorate like in Section \ref{lowering}. The results are shown in Figure \ref{fig:distance_constant_1E-10_constant_kappa}.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[trim=0.2cm 0cm 0cm 1.35cm,width=15cm, clip]{../picture/New_model/1E-10/distance_constant_1E-10.png}
		\end{center}
		\caption{Maximum distance (infinity norm) of the steady-state probabilities from the reference high-precision ACTMC ($10^{-20}$) against various values of $k$ for lowered absolute termination epsilon $\epsilon = 10^{-10}$. ACTMC results for $\epsilon = 10^{-10}$ are also compared to the reference.}
		\label{fig:distance_constant_1E-10_constant_kappa}
	\end{figure}

	%\pagebreak
	\subsection{Adjusting epsilon $\epsilon = 10^{-10}$ to $k$}
	\label{adjusting}

	Since increasing $k$ increases the amount of phases (intermediate states) and each phase may have error up to $\epsilon$, the actual potential error for each state of the queue is $\epsilon \cdot k$. To compensate for this, we devise a simple formula
	$$\epsilon^{'} = \frac{\epsilon}{k}$$
	where $\epsilon^{'}$ is the adjusted epsilon that should be given to PRISM if precision $\epsilon$ is desired for $k$ phases.
	Using adjusted epsilon, the results get significantly better, as shown in Figure \ref{fig:distance_dynamic_1E-10_constant_kappa}. What's more, the phase-type fitting results now seem to converge towards the reference steadily as $k$ increases. This indicates that this approach is correct, and that the reference is trustworthy.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[trim=0.2cm 0cm 0cm 1.35cm,width=15cm, clip]{../picture/New_model/1E-10/distance_dynamic_1E-10.png}
		\end{center}
		\caption{Maximum distance (infinity norm) of the steady-state probabilities from the reference high-precision ACTMC ($10^{-20}$) against various values of $k$ for lowered and adjusted absolute termination epsilon $\epsilon = 10^{-10}$. ACTMC results for $\epsilon = 10^{-10}$ are also compared to the reference.}
		\label{fig:distance_dynamic_1E-10_constant_kappa}
	\end{figure}

	\subsection{Required computation time to obtain precise results}
	\label{computation_time}

	Although we have shown it is possible to obtain relatively precise results using phase-type distributions, the precision comes at a rather steep cost in computation time. This is shown in Figure \ref{fig:performance_dynamic}. Note that even though Storm is one of the fastest available tools, it still struggles to deliver good precision within reasonable time. The computations all ran on the same virtual machine under same conditions\footnote{Running in a virtual machine might have reduced the performance. The virtual machine we used is available at the \href{http://www.stormchecker.org/documentation/vm/vm.html}{Storm website}.}.

	\begin{figure}
		\begin{center}
			\includegraphics[trim=0.2cm 0cm 0cm 1.35cm,width=15cm, clip]{../picture/New_model/1E-6/performance_dynamic.png}
		\end{center}
		\caption{Time of computation in seconds of the steady-state probabilities against various values of $k$ for lowered and adjusted absolute termination epsilon $\epsilon = 10^{-10}$, i.e. corresponding to the data shown in Figure \ref{fig:distance_dynamic_1E-10_constant_kappa}. The computation of our PRISM ACTMC is constant in regards to $k$ because it does not use phase-type fitting. }
		\label{fig:performance_dynamic}
	\end{figure}

	\subsection{Another Model : firstRejuvenation}

	We can observe similar results in terms of lack of precision on another model Figure \ref{fig:fr_1E-10_constant}. The precision must also depend on the PH fitting parameter k in this model. Here we went through difficulties related to the methods used by the engine in Prism. The Power method for computing SSP uses different iterations which are matrix multiplication. The number of iteration will strongly depend on the model and its parameter like timeouts and rates. In the First Rejuvenation model The convergence for the power method is really slow and the number of iteration required is important. In the queue, 100 000 000 iterations was enough to compute the SSP, but here the number of iteration prevent us to compute the SSP for k greater than 1000 with reasonable precision.

	The results of computation are given in the Figure \ref{fig:fr_1E-10_constant}. The orange curve is the error versus the number of phase $k$. The precision setting of the computation is $10^{-10}$ but the error is much bigger (at least $10^{-3}$). This error can be related to the theoretical approximation error of the PH fitting, but also the error of computation of the engine.
	\begin{figure}[H]
		\begin{center}
			\includegraphics[trim=0.2cm 0cm 0cm 1.35cm,width=15cm, clip]{../picture/firstRejuvenation/distance_1E-10_constant.png}
		\end{center}
		\caption{Maximum distance (infinity norm) of the steady-state probabilities from the reference high-precision ACTMC ($10^{-20}$) against various values of $k$ for lowered and adjusted absolute termination epsilon $\epsilon = 10^{-10}$. ACTMC results for $\epsilon = 10^{-10}$ are also compared to the reference.}
		\label{fig:fr_1E-10_constant}
	\end{figure}

	%\pagebreak
	\section{Conclusion}

%	CHANGE THIS

	We have presented a reliable method for high-precision analysis of PRISM CTMC models with phase-type distributions, that is, adjusting the termination epsilon by the number of phases. However, the large number of phases drastically increases the computation time.

	Using the default setting of precision when doing PH fitting is not reliable at all. The relation between the experimental precision and the parameter of the model is not trivial. We have presented a sufficient condition to increased the precision while increasing the number of phases.

	For better performance, we suggest usage of more specialized tools that can deal with non-Markovian distributions directly, without phase-type fitting. Our PRISM ACTMC extension \footnote{Our PRISM ACTMC extension is not fully completed yet and lacks optimization. It will be officially released later, but early development version is available at \href{https://github.com/VojtechRehak/prism-gsmp}{https://github.com/VojtechRehak/prism-gsmp}.} has delivered results with sufficient precision a lot faster.

\iffalse
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.
\section{Conclusion}
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}

\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017

\end{thebibliography}
\fi
\printbibliography
\end{document}